Data Specialization for a Staged Lambda Calculus

### 2014-11-07 paper/splitting.tex
Editing pass through stage 1 terminals.

Splitting a stage \bbone\ term in \lang\
% $\colone e A$ 
(where $e\mathbin{\redonesym}[\xi;v]$), yields a term 
$e_1$ that performs all stage~\bbone\ subcomputations in $e$ and a \emph{resumer} function $e_2 = l.r$ that performs all stage~\bbtwo\ subcomputations.
For splitting to match the results of evaluating $e$ in \lang,
we require that $e_1$ reduce to the tuple $(y,z)$,
where $y$ is identical to $\masko{\xi;v}$, and 
$\letin l z r$ reduces to the same value as $\maskt{\xi;v}$. Intuitively, $e_1$ performs both the stage~\bbone\ subcomputations needed to produce an \emph{immediate result} $y$, and the stage~\bbone\ subcomputations (called \emph{precomputations}) required to produce the input required by the resumer.

The splitting algorithm for stage~\bbone\ terms, as specified by the
judgment $e \splitonesym [e_1,l.r]$ in \cref{fig:termSplitOne}, proceeds
recursively on the structure of~$e$.

When $e$ is a terminal (variable or base constant), the combined result
is formed by tupling $e$ with a \texttt{()} precomputation, 
and the resumer is just constant \texttt{()}.  
For example, \texttt{3} simply splits into combined result \texttt{((3,())} and \texttt{\_=>4}.
When $e$ is a non-terminal (except $\next$),
splitting descends into $e$, recursively splitting subterms
and generating combined terms $c_1,\ldots,c_n$.
The new combined term of $e$ then works by binding $c_1,\ldots,c_n$
to the patterns $(y_1,z_1),\ldots,(y_n,z_n)$
to isolate their immediate results and precomputations.
The immediate result of $e$ is then formed by replacing its subterms with $y_1,\ldots,y_n$, 
and the precomputation is formed from the tuple of $z_1,\ldots,z_n$.

>>>>

Splitting a stage \bbone\ term in \lang\
% $\colone e A$ 
(where $e\mathbin{\redonesym}[\xi;v]$), yields a term 
$e_1$ that performs all stage~\bbone\ subcomputations in $e$ and a \emph{resumer} function $e_2 = l.r$ that performs all stage~\bbtwo\ subcomputations.
To ensure the results of splitting produce the same output as the evaluation $e$ in \lang,
we require that $e_1$ reduce to the tuple $(y,z)$,
where $y$ is identical to $\masko{\xi;v}$, and 
$\letin l z r$ reduces to the same value as $\maskt{\xi;v}$. Since $e_1$ performs both stage~\bbone\ subcomputations needed to produce an \emph{immediate result} $y$, and the stage~\bbone\ subcomputations (called \emph{precomputations}) required to produce the input required by the resumer, we refer to this term as the \emph{combined result}, and hereafter denote it as $c$.

The splitting algorithm for stage~\bbone\ terms, as specified by the
judgment $e \splitonesym [c,l.r]$ in \cref{fig:termSplitOne}, proceeds
recursively on the structure of~$e$.
When $e$ is a terminal (a variables or base constant)
splitting yields a combined result formed by tupling $e$ with a \texttt{()} precomputation, and the trivial resumer \texttt{()}. (Stage~\bbone\ terminals, by definition, contain no stage~\bbtwo\ content.)  For example, \texttt{3} splits into combined result \texttt{(3,())} and resumer \texttt{\_=>()}.

% Note the direct correspondence of these splitting outputs to the output of the stage~\bbone\ and~\bbtwo\ masking functions.

When $e$ is a non-terminal (except $\next$),
splitting descends into $e$, recursively splitting subterms
and generating combined terms $c_1,\ldots,c_n$.
The new combined term of $e$ then works by binding $c_1,\ldots,c_n$
to the patterns $(y_1,z_1),\ldots,(y_n,z_n)$
to isolate their immediate results and precomputations.
The immediate result of $e$ is then formed by replacing its subterms with $y_1,\ldots,y_n$, 
and the precomputation is formed from the tuple of $z_1,\ldots,z_n$.
